% !TeX root = main.tex

\graphicspath{ {images/} }

\titledquestion{Analyzing NMT Systems}[25]

\begin{parts}

    \part[3] Look at the {\monofam{src.vocab}} file for some examples of phrases and words in the source language vocabulary. When encoding an input Mandarin Chinese sequence into ``pieces'' in the vocabulary, the tokenizer maps the sequence to a series of vocabulary items, each consisting of one or more characters (thanks to the {\monofam{sentencepiece}} tokenizer, we can perform this segmentation even when the original text has no white space). Given this information, how could adding a 1D Convolutional layer after the embedding layer and before passing the embeddings into the bidirectional encoder help our NMT system? \textbf{Hint:} each Mandarin Chinese character is either an entire word or a morpheme in a word. Look up the meanings of 电, 脑, and 电脑 separately for an example. The characters 电 (electricity) and  脑 (brain) when combined into the phrase 电脑 mean computer.

    \ifans{Since Chinese text often lacks explicit word boundaries, the convolutional layer can identify meaningful character combinations, helping the model learn relevant information and improving translation quality by better handling the intricacies of the language's character-based structure.}


    \part[8] Here we present a series of errors we found in the outputs of our NMT model (which is the same as the one you just trained). For each example of a reference (i.e., `gold') English translation, and NMT (i.e., `model') English translation, please:
    
    \begin{enumerate}
        \item Identify the error in the NMT translation.
        \item Provide possible reason(s) why the model may have made the error (either due to a specific linguistic construct or a specific model limitation).
        \item Describe one possible way we might alter the NMT system to fix the observed error. There are more than one possible fixes for an error. For example, it could be tweaking the size of the hidden layers or changing the attention mechanism.
    \end{enumerate}
    
    Below are the translations that you should analyze as described above. Only analyze the underlined error in each sentence. {\bf Rest assured that you don't need to know Mandarin to answer these questions. You just need to know English!} If, however, you would like some additional color on the source sentences, feel free to use a resource like \url{https://www.archchinese.com/chinese_english_dictionary.html} to look up words. Feel free to search the training data file to have a better sense of how often certain characters occur.

    \begin{subparts}
        \subpart[2]
        \textbf{Source Sentence:} 贼人其后被警方拘捕及被判处盗窃罪名成立。 \newline
        \textbf{Reference Translation:} \textit{\underline{the culprits were} subsequently arrested and convicted.}\newline
        \textbf{NMT Translation:} \textit{\underline{the culprit was} subsequently arrested and sentenced to theft.}
        
        \ifans{\\
        	1. The number agreement is incorrect. \\
        	2. Chinese does not use explicit plural forms as in English.\\
        	3. Enhancing the model's understanding of plurality and agreement in the context of Chinese sentences could help address this error. This might involve training the model on a more diverse dataset with a focus on agreement rules and structures in both languages.
        	}


        \subpart[2]
        \textbf{Source Sentence}: 几乎已经没有地方容纳这些人,资源已经用尽。\newline
        \textbf{Reference Translation}: \textit{there is almost no space to accommodate these people, and resources have run out.   }\newline
        \textbf{NMT Translation}: \textit{the resources have been exhausted and \underline{resources have been exhausted}.}
        
        \ifans{\\
        1. The NMT translation repeats "resources have been exhausted" twice, leading to redundancy. \\
        2. The model may have generated redundant text due to a lack of contextual understanding. \\
        3. Implementing better redundancy detection mechanisms in the NMT system could help prevent such errors.
        }

        \subpart[2]
        \textbf{Source Sentence}: 当局已经宣布今天是国殇日。 \newline
        \textbf{Reference Translation}: \textit{authorities have announced \underline{a national mourning today.}}\newline
        \textbf{NMT Translation}: \textit{the administration has announced \underline{today's day.}}
        
        \ifans{\\
        1. The NMT translation incorrectly translates "today's day," which is redundant. \\
        2.The model might not have sufficient context or knowledge about cultural and linguistic nuances.\\
        3. One way to address this error is to enrich the training data with more examples of similar terms and their translations in the target language.
        }
        
        \subpart[2] 
        \textbf{Source Sentence\footnote{This is a Cantonese sentence! The data used in this assignment comes from GALE Phase 3, which is a compilation of news written in simplified Chinese from various sources scraped from the internet along with their translations. For more details, see \url{https://catalog.ldc.upenn.edu/LDC2017T02}. }:} 俗语有云:``唔做唔错"。\newline
        \textbf{Reference Translation:} \textit{\underline{`` act not, err not "}, so a saying goes.}\newline
        \textbf{NMT Translation:} \textit{as the saying goes, \underline{`` it's not wrong. "}}
        
        \ifans{\\
        1. The translation fails to capture the meaning of the source idiom. Instead, it provides a literal translation that does not convey the intended message. \\
        2. The model may lack sufficient understanding of idiomatic expressions and their cultural nuances, resulting in a literal translation.\\
        3. Training the NMT model with a broader range of idiomatic expressions could help improve the translation of idioms.
        }
    \end{subparts}


    \part[14] BLEU score is the most commonly used automatic evaluation metric for NMT systems. It is usually calculated across the entire test set, but here we will consider BLEU defined for a single example.\footnote{This definition of sentence-level BLEU score matches the \texttt{sentence\_bleu()} function in the \texttt{nltk} Python package. Note that the NLTK function is sensitive to capitalization. In this question, all text is lowercased, so capitalization is irrelevant. \\ \url{http://www.nltk.org/api/nltk.translate.html\#nltk.translate.bleu_score.sentence_bleu}
    } 
    Suppose we have a source sentence $\bs$, a set of $k$ reference translations $\br_1,\dots,\br_k$, and a candidate translation $\bc$. To compute the BLEU score of $\bc$, we first compute the \textit{modified $n$-gram precision} $p_n$ of $\bc$, for each of $n=1,2,3,4$, where $n$ is the $n$ in \href{https://en.wikipedia.org/wiki/N-gram}{n-gram}:
    \begin{align}
        p_n = \frac{ \displaystyle \sum_{\text{ngram} \in \bc} \min \bigg( \max_{i=1,\dots,k} \text{Count}_{\br_i}(\text{ngram}), \enspace \text{Count}_{\bc}(\text{ngram}) \bigg) }{\displaystyle \sum_{\text{ngram}\in \bc} \text{Count}_{\bc}(\text{ngram})}
    \end{align}
     Here, for each of the $n$-grams that appear in the candidate translation $\bc$, we count the maximum number of times it appears in any one reference translation, capped by the number of times it appears in $\bc$ (this is the numerator). We divide this by the number of $n$-grams in $\bc$ (denominator). \newline 

    Next, we compute the \textit{brevity penalty} BP. Let $len(c)$ be the length of $\bc$ and let $len(r)$ be the length of the reference translation that is closest to $len(c)$ (in the case of two equally-close reference translation lengths, choose $len(r)$ as the shorter one). 
    \begin{align}
        BP = 
        \begin{cases}
            1 & \text{if } len(c) \ge len(r) \\
            \exp \big( 1 - \frac{len(r)}{len(c)} \big) & \text{otherwise}
        \end{cases}
    \end{align}
    Lastly, the BLEU score for candidate $\bc$ with respect to $\br_1,\dots,\br_k$ is:
    \begin{align}
        BLEU = BP \times \exp \Big( \sum_{n=1}^4 \lambda_n \log p_n \Big)
    \end{align}
    where $\lambda_1,\lambda_2,\lambda_3,\lambda_4$ are weights that sum to 1. The $\log$ here is natural log.
    \newline
    \begin{subparts}
        \subpart[5] Please consider this example: \newline
        Source Sentence $\bs$: \textbf{需要有充足和可预测的资源。} 
        \newline
        Reference Translation $\br_1$: \textit{resources have to be sufficient and they have to be predictable}
        \newline
        Reference Translation $\br_2$: \textit{adequate and predictable resources are required}
        
        NMT Translation $\bc_1$: there is a need for adequate and predictable resources
        
        NMT Translation $\bc_2$: resources be suﬀicient and predictable to
        
        Please compute the BLEU scores for $\bc_1$ and $\bc_2$. Let $\lambda_i=0.5$ for $i\in\{1,2\}$ and $\lambda_i=0$ for $i\in\{3,4\}$ (\textbf{this means we ignore 3-grams and 4-grams}, i.e., don't compute $p_3$ or $p_4$). When computing BLEU scores, show your work (i.e., show your computed values for $p_1$, $p_2$, $len(c)$, $len(r)$ and $BP$). Note that the BLEU scores can be expressed between 0 and 1 or between 0 and 100. The code is using the 0 to 100 scale while in this question we are using the \textbf{0 to 1} scale. Please round your responses to 3 decimal places. 
        \newline
        
        Which of the two NMT translations is considered the better translation according to the BLEU Score? Do you agree that it is the better translation?
        
        \ifans{
        	Let me consider $\bc_1$. It has $\text{len}(\bc_1) = 9$ unigrams "there", "is", "a", "need", "for", "adequate", "and", "predictable", "resources". Unigrams "adequate" and "and" appears in $\br_2$, while unigrams "predictable" and "resources" appear in both $\br_1$ and $\br_2$, hence
        	\begin{align}
        		p_1(\bc_1) = \frac{1+1+1+1}{9} = \frac{4}{9}.
        	\end{align}
        	Next, there are $\text{len}(\bc_1) - 1 = 8$ bigrams in $\bc_1$. It shares the following three bigrams with $\br_2$: "adequate and", "and predictable", "predictable resources".
        	\begin{align}
        		p_2(\bc_1) = \frac{1+1+1}{8} = \frac{3}{8}.
        	\end{align}
        	
        	Let me consider $\bc_2$. It has $\text{len}(\bc_2) = 6$ unigrams and $\text{len}(\bc_1) - 1 = 5$ bigrams.
        	\begin{align}
        		p_1(\bc_2) &= \frac{6}{6} = 1,	&
        		p_2(\bc_2) = \frac{3}{5} = \frac{3}{5}.
        	\end{align}
        	
        	Brevity penalty. First of all, $\text{len}(\br_1) = 11$, $\text{len}(\br_2) = 6$, $\text{len}(\bc_1) = 9$, $\text{len}(\bc_2) = 6$. This means that $\text{len}(\br | \bc_1) = 11$ and $\text{len}(\br | \bc_2) = 6$. Hence $\text{BP}(\bc_1) = \exp(-2/9)$ and $\text{BP}(\bc_2) = 1$.
        	
        	Finally,
        	\begin{align}
        		\text{BLEU}(\bc_1) = \exp\left(-\frac{2}{9} + \frac12 \ln \frac{1}{3} + \frac12 \ln \frac{3}{8} \right) = .327
        	\end{align}
        	\begin{align}
        		\text{BLEU}(\bc_2) = \exp\left( 0 + \frac12 \ln \frac{3}{5} \right) = \sqrt{\frac{3}{5}} = .775
        	\end{align}
        	
        	Translation $\bc_2$ has higher BLEU score, but it's obviously worst translation, it's grammaticlaly incorrect.
        }
        
        \subpart[5] Our hard drive was corrupted and we lost Reference Translation $\br_1$. Please recompute BLEU scores for $\bc_1$ and $\bc_2$, this time with respect to $\br_2$ only. Which of the two NMT translations now receives the higher BLEU score? Do you agree that it is the better translation?
        
        \ifans{
        	Let's compute BLEU with respect of $\br_2$ only. Common unigrams between $\bc_1$ and $\br_2$ are: "adequate", "and", "predictable", "resources".
        	\begin{align}
        		p_1(\bc_1) &= \frac{4}{9},	&
        		p_2(\bc_1) = \frac{3}{8}
        	\end{align}
        	Common unigrams between $\bc_2$ and $\br_2$ are: "and", "predictable", "resources".
        	\begin{align}
        		p_1(\bc_2) &= \frac{3}{6} = \frac{1}{2},	&
        		p_2(\bc_2) = \frac{1}{5}.
        	\end{align}
        	Brevity penalties are absent $\text{BP}(\bc_1) = 1$ and $\text{BP}(\bc_2) = 1$.
        	
        	Finally
        	\begin{align}
        		\text{BLEU}(\bc_1) = \exp\left( \frac12 \ln \frac{4}{9} + \frac12 \ln \frac{3}{8} \right) = \frac{1}{\sqrt{6}} = .408
        	\end{align}
        	\begin{align}
        		\text{BLEU}(\bc_2) = \exp\left( \frac12 \ln \frac{1}{2} + \frac12 \ln \frac{1}{5} \right) = \frac{1}{\sqrt{10}} = .316
        	\end{align}
        	Translation $\bc_1$ has higher BLEU score, I agree that $\bc_1$ is better translation.
        }
        
        \subpart[2] Due to data availability, NMT systems are often evaluated with respect to only a single reference translation. Please explain (in a few sentences) why this may be problematic. In your explanation, discuss how the BLEU score metric assesses the quality of NMT translations when there are multiple reference transitions versus a single reference translation.
        
        \ifans{
        	Evaluating NMT systems with only a single reference translation is problematic because it ignores the natural diversity of language—there are often many correct ways to translate the same text. The BLEU score, when based on a single reference, may unfairly penalize valid translations that diverge from that reference.
        	With multiple references, BLEU can more accurately assess translation quality by recognizing a broader range of acceptable translations, thus offering a more comprehensive evaluation of an NMT system's performance by acknowledging linguistic variability.
        }
        
        \subpart[2] List two advantages and two disadvantages of BLEU, compared to human evaluation, as an evaluation metric for Machine Translation. 
        
        \ifans{
        	Advantages of BLEU:
        	
        	1. Scalability: BLEU allows for the automatic and rapid evaluation.
        	
        	2. Objectivity: It provides a consistent and objective metric for evaluation.
        	
        	
        	Disadvantages of BLEU:
        	
        	1. Lack of Semantic Understanding: BLEU may not fully capture the semantic accuracy or the fluency of the translation.
        	
        	2. Dependence on Reference Quality.
        }
        
    \end{subparts}
\end{parts}
